
---
title: "Introduction to fluorojip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fluorojip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(fluorojip)
```

## Overview

`fluorojip` computes **chlorophyll *a* fluorescence OJIP parameters** from raw fluorescence data.  
The package focuses on **PSII performance** and **JIP-test indices**, following standard biophysical equations.

This vignette presents a basic workflow to:

1. Load raw OJIP fluorescence data (example dataset or Handy PEA files)
2. Compute JIP-test parameters
3. Visualize results using **heatmaps** and **3D surface plots**

---

## Example Dataset

The package includes an example dataset containing typical fluorescence values recorded with a fluorometer.

```{r}
data(example_fluorojip)
head(example_fluorojip)
```

---

## Computing OJIP Parameters

Use `calc_fluorojip()` to compute the full suite of JIP-test indices.  
The function automatically handles intermediate variables such as  
$V_j$, $V_i$, and $M_0$, ensuring scientific consistency.

```{r}
res <- calc_fluorojip(example_fluorojip)
head(res)
```

### Key parameters calculated

**Yields**
- `phi_Po` ($F_v/F_m$)
- `psi_Eo`
- `phi_Eo`

**Fluxes per reaction center (RC)**
- `ABS_RC`
- `TRo_RC`
- `ETo_RC`
- `DIo_RC`

**Fluxes per cross section (CS)**
- `ABS_CSm`
- `TRo_CSm`
- `ETo_CSm`
- `DIo_CSm`

**Performance**
- `PI_abs` (Performance Index)

---

## Multivariate Visualization

When many parameters and treatments are involved, `fluorojip` provides modern multivariate alternatives to traditional univariate charts.

### Heatmaps

Heatmaps allow simultaneous visualization of patterns across parameters and treatments.  
Data are automatically scaled (Z-score) to make variables with different units comparable.

```{r}
# Selecting key parameters for visualization
params <- c("phi_Po", "PI_abs", "ABS_RC", "DIo_RC", "ETo_CSm")

plot_param_heatmap(
  res,
  group_col = "treatment",
  params    = params,
  scale     = "zscore",
  main      = "JIP-Test Parameters Heatmap"
)
```

### 3D Surface Plots

3D surface plots help visualize the overall "data landscape", highlighting treatments
with the strongest effects on PSII performance.

```{r}
plot_3d_fluorojip(res, params = c("Fv_Fm", "PI_abs", "area"))
```

---

## Reading Handy PEA Data

If your fluorescence data were exported directly from a **Handy PEA** fluorometer,
you can use the built-in reader:

```{r}
# my_data <- read_handypea_csv("path/to/your_data.csv")
# res_pea <- calc_fluorojip(my_data)
```

---

## Summary

The `fluorojip` workflow provides a **reproducible path** from raw fluorescence measurements
to **publication-ready multivariate visualizations**, ensuring that all calculations strictly
follow the standard JIP-test theory described in the current literature.
